### 2019-06-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
A la BU le samedi après-midi ouais
Almost done for ch5

We consider $M \leq K$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
A perfect time and frequency synchronization is assumed,
and some \iid{} background traffic is assumed to be non-uniformly in the $K$ channels.
As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent uplink packet can be received by the gateway.
%
Unfortunately, it is very hard to formally analyze the IoT network models we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic (\ie, learning) devices.
Even if there are many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step the number of communicating devices is random and unpredictable.

@@

% Another interesting extension of the studied model is to consider different channels utility for each devices. It is a very interesting extension of our model, as for instance the devices could be located on different parts of a building or a field and suffer from different mean qualities of access to each channels.
% This extension is not studied per say but we quickly review the existing literature, consisting in the two recent works \cite{Bistritz18,KaufmannAbbas19}, and we discuss the possible real-world usage of the existing algorithms.

@@

Note that when $n$ players choose arm $k$ at time $t$, this counts as $n$ collisions, not just one. So $\cC_k(T)$ counts the total \emph{number of colliding players} rather than the number of collision events. Hence there is small abuse of notation when calling it a number of collisions.

Letting $\cP_M = \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
be the set of bandit instances such that there is a strict gap between the $M$ best arms and the other arms, we now provide a regret decomposition for any $\boldsymbol{\mu} \in \cP_M$.

@@

Similarly to what we present above in Section~\ref{sec:2:lowerUpperBoundsRegret},
we use the Kullback-Leibler divergence to express our lower bound.
Let $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$ be KL divergence between the Bernoulli distribution of mean $x$ and that of mean $y$.
%
We first introduce the assumption under which we derive a regret lower bound, that generalizes the assumption of uniform efficiency given in Definition~\ref{def:2:uniformlyEfficientAlgorithm},
a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
% Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
% This class contains our proposals, \RandTopM{} and \MCTopM, but rules out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.

@@

Having a small regret on every problem instance, \ie, being uniformly efficient,
is a natural assumption for algorithms,
that rules out algorithms tuned to perform well on specific instances only.
%
From this assumption $\left(R_T(\boldsymbol{\mu},M,\cA)=o(T^\alpha)\right)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
${T}- \E_{\mu}[T_k(T)] {=} o(T^{\alpha})$, and so
\begin{equation}\label{eq:5:intermediate}
  % \vspace*{-5pt}  % XXX
  \sum_{j =1}^M\left(\frac{T}{M} - \E_{\mu}[T^j_k(T)]\right) = o(T^{\alpha}).
  % \vspace*{-5pt}  % XXX
\end{equation}

The additional assumption in \eqref{eq:5:SUE} further implies some notion of \emph{fairness}, as it suggests that each of the $M$ players spends on average the same amount of time on each of the $M$ best arms. Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\cA_{\sigma} = (\cA_{\sigma(1)},\dots,\cA_{\sigma(M)})$ is independent from the choice of permutation $\sigma \in \Sigma_M$. In that case, it holds that  $\E_{\mu}[T_k^j(T)]=\E_{\mu}[T_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in [M]$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency. Note that all our proposed algorithms are permutation invariant and \MCTopM{} is thus an example of strongly uniformly efficient algorithm, as we prove in Section~\ref{sec:5:upperbounds} that its regret is logarithmic on every instance $\mu \in \cP_M$.

As we state it here, the constraint on fairness is new, and it was not studied as such in previous works.
In the proofs of the regret bounds for \rhoRand{} in \cite{Anandkumar11}, the authors briefly mention that their policy is invariant under permutation and that this yields a certain fairness guarantee, but without formalizing it more.
We note that the notion of fairness we defined above can be interpreted as a ``cooperative fairness'', opposed to the notion of ``arm fairness'' that has been studied in a few papers on single-player stochastic MAB.
For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of the total time steps, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K]$ is given and known to the algorithm.

@@

The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
Fix a player $j\in[M]$, and we consider the observations $O_t$ gathered by this player after $t$ rounds of its algorithm $\cA_j$:
\[ O_t := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right),\]
where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).

Fix a problem $\bm{\mu}$,
% and as we do in Appendix~\ref{proof:5:BetterLowerBound},
then introduce an alternative model parametrized by $\bm{\lambda}$, with a small difference between $\bm{\mu}$.
Fix a sub-optimal arm $k$ and $\varepsilon>0$, and $\lambda_k = \mu^*_M + \varepsilon$ and $\lambda_{\ell} = \mu_{\ell}$ for any $\ell\neq k$.
We denote $\Pr_{\bm{\mu}}^{O_t}$ the distribution of the vector $O_t$ under the model $\bm{\mu}$ when using algorithm $\cA$.
Then we introduce the \emph{collision information term} as:
\begin{equation}\label{eq:5:collisionInformationTerm}
  \cI_{\bm{\mu},\bm{\lambda}}(\cA,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}}).
\end{equation}

For a \emph{decentralized strategy} $\cA$ that has access to the sensing information (\ie, ruling out model \modeltrois), and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T) = \smallO{\log(T)}$,
we now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections.
The additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
The theorem stated below is proven in the Appendix of \cite{Besson2018ALT},
% \ref{proof:5:BetterLowerBound},
and it also yields an asymptotic logarithmic lower bound on the regret.

\begin{theorem}\label{thm:5:BetterLowerBound}
\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
  Under observation models $\modelun$ and $\modeldeux$, for any strongly uniformly efficient \emph{decentralized} policy $\cA$ and $\boldsymbol{\mu}\in\cP_M$.
  Furthermore, if $\cA$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T) = \smallO{\log(T)}$, then
  % \vspace*{-10pt}  % XXX
  \begin{equation}\label{eq:5:LBDraws}
    \forall j \in [M], \ \forall k \in \Mworst, \ \ \ \liminf_{T\to \infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k, \mu_M^*)}.
  \end{equation}

  % \vspace{-0.3cm}  % XXX
  \noindent From Lemma~\ref{lem:5:1stLowerBound}, it follows that
  % \vspace*{-5pt}  % XXX
  \begin{equation}\label{eq:5:ourLowerBound}
    \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log(T)}
    \geq M \times \left( \sum_{k \in \Mworst} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_M^*)} \right) .
  \end{equation}
\end{leftbar}  % XXX leftbar theorembar, comment if needed
\end{theorem}

Observe that the regret lower bound \eqref{eq:5:ourLowerBound} is tighter than the state-of-the-art lower bound in this setup
given by \cite{Zhao10}, that states that
\begin{equation}\label{eq:5:Zhao10LowerBound}
  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log(T)}
  \geq \sum_{k \in \Mworst} \left( \sum_{j=1}^{M} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_{j}^*)} \right),
\end{equation}
as for every $k \in \Mworst$ and $j \in [M]$, $\kl(\mu_k, \mu_j^*) \geq \kl(\mu_k, \mu_M^*)$
(see Figure~\ref{fig:5:CompLowerBounds} in Appendix~\ref{app:5:illustrationLowerBound}).
%
It is worth mentioning that \cite{Zhao10} proved a lower bound under the more general assumption for $\cA$ that there exists some numbers $(a_{k}^j)$ such that $a_{k}^j T - \E_{\mu}[T_k^j(T)] = o(T^\alpha)$ whereas in Definition~\ref{def:5:DecentralizedUniformEfficiency} we make the choice $a_{k}^j = 1/M$.
Our result could be extended to this case but we chose to keep the notation simple and focus on \emph{fair allocation} of the optimal arms between players.

@@

Under observation model \modeltrois{} no sensing information is available and the previous algorithms cannot be used, as the sum of sensing information $S_k^j(t)$ and thus the empirical mean $\widehat{\mu}_k^j(t)$ cannot be computed, hence neither the indices $U_k^j(t)$. However, one can still define a notion of \emph{empirical reward} received from arm $k$ by player $j$, by introducing
%
% \vspace*{-5pt}
\begin{equation}
  \widetilde{S_k}^j(t) = \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
  \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) := \widetilde{S_k}^j(t) \;/\; T_k^j(t).
\end{equation}

Note that $\widetilde{\mu_k}^j(t)$ is no longer meant to be an unbiased estimate of $\mu_k$ as it also takes into account the collision information, that is present in the reward. Based on this empirical reward, one can similarly defined modified indices as
%
\begin{equation}\label{eq:5:indexTilde}
  \widetilde{U_k}^j(t) = \begin{cases}
      \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widetilde{\mu_k}^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB.
  \end{cases}
\end{equation}

Given any of these two index policies (\UCB{} or \klUCB), the \Selfish{} algorithm is defined by,
%
\begin{equation}
  A^j(t) = \argmax_{ k \in [K]} \ \widetilde{U_k}^j(t-1).\label{algo:5:Selfish}
\end{equation}
The name comes from the fact that each player is targeting, in a ``selfish'' way, the arm that has the highest index, instead of accepting to target only one of the $M$ best.
The reason that this may work precisely comes from the fact that $\widetilde{U_k}^j(t)$ is no longer an upper-confidence on $\mu_k$,
but some hybrid index that simultaneously increases when a transmission occurs and decreases when a collision occurs.

This behavior is easier to be understood for the case of \Selfish-\UCB{} in which, letting $N_k^{j,C}(t) = \sum_{s=1}^t \indic(C^j(t))$ be the number of collisions on arm $k$, one can show that the hybrid \Selfish{} index induces a penalty proportional to the fraction of collision on this arm and the quality of the arm itself:
\begin{equation}
  \widetilde{U_k}^j(t) = U_k^j(t) -
  \underbrace{\left(\frac{N_k^{j,C}(t)}{N_k^j(t)}\right)}_{\text{fraction of collisions}}
  \underbrace{\left(\frac{1}{N_k^{j,C}(t)}\sum_{t=1}^{T}Y_{A^j(t),t} \indic(C^j(t)) \indic(A^j(t) = k)\right)}_{\text{estimate of } \mu_k }.
\end{equation}

@@

We illustrate in Section~\ref{sub:5:SelfishFails} that \Selfish{} does have a (very) small probability to fail (badly), for some problem with small $K$,
which precludes the possibility of a logarithmic regret for any problem.
In most cases, it empirically performs similarly to all the algorithms described before,
and usually outperforms \rhoRand,
even if it neither exploits the sensing information, nor the knowledge of the number of players $M$.
%
Practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or cannot be considered.
Thus we propose next our main contribution, the \MCTopM{} algorithm, proved to be asymptotically optimal for the identification of suboptimal arms (when using \klUCB), attaining order-optimal logarithmic regret, and outperforming all the other algorithms for the ``sensing case''.

@@

In a single-player setting, \emph{index policies} are popular bandit algorithms: at each round one index is computed for each arm, that only depends on the history of plays of this arm and (possibly) some exogenous randomness. Then, the arm with highest index is selected. This class of algorithms includes the UCB family, in which the index of each arm is an Upper Confidence Bound for its mean, but also some Bayesian algorithms like Bayes-UCB \citep{Kaufmann12BUCB} or the randomized Thompson Sampling algorithm \citep{Thompson33,AgrawalGoyal11,Kaufmann12Thompson}.

% concrete examples and how their are used within MPB

\paragraph{Index policies.}
%
The approaches we now describe for multi-player bandits can be used in combination with any index policy, but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained. In particular, we focus on two types of indices:
\UCB{} indices \citep{Auer02}
and \klUCB{} indices \citep{KLUCBJournal}, that can be defined for each player $j$ in the following way.
%
Letting $S_k^j(t) := \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained by player $j$ for arm $k$,
then $\widehat{\mu}_k^j(t) = S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$ for player $j$, and thus one can define the index
\begin{equation}\label{eq:5:indexFor_UCB_klUCB}
  U_k^j(t) := \begin{cases}
      \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widehat{\mu}_k^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB,
  \end{cases}
\end{equation}
where $f(t)$ is some \emph{exploration function}. $f(t)$ is usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
A classical (single-player) UCB algorithm aims at the arm with largest index. However, if each of the $M$ players selects the arm with largest UCB, all the players will end up colliding most of the time on the best arm.
To circumvent this problem, several coordination mechanisms have emerged, that rely on \emph{ordering} the indices and targeting \emph{one of} the $M$-best indices.

>>>>

We consider $M$ identical dynamic IoT devices, communicating with a unique gateway, in $K$ orthogonal channels and in a acknowledgment-based wireless protocol slotted in time.
A perfect time and frequency synchronization is assumed,
and some \iid{} background traffic is assumed to be non-uniformly in the $K$ channels.
As before, if two or more devices decide to use the same channel at the same time, a \emph{collision} arises and none of sent uplink packet can be received by the gateway.
%
Unfortunately, it is very hard to formally analyze the IoT network models we presented in Chapter~\ref{chapter:4}, mainly because of the random activation process of all the dynamic (\ie, learning) devices.
Even if there are many identical bandit algorithms learning independently and in a decentralized way, the difficulty mainly comes from the fact that all of them are only communicating at some (random) time steps, and at each time step the number of communicating devices is random and unpredictable.

@@

% Another interesting extension of the studied model is to consider different channels utility for each device. It is a very interesting extension of our model, as for instance the devices could be located on different parts of a building or a field and suffer from different mean qualities of access to each channel.
% This extension is not studied per say but we quickly review the existing literature, consisting in the two recent works \cite{Bistritz18,KaufmannAbbas19}, and we discuss the possible real-world usage of the existing algorithms.

@@

Note that when $n$ players choose arm $k$ at time $t$, this counts as $n$ collisions, not just one. So $\cC_k(T)$ counts the total \emph{number of colliding players} rather than the number of collision events. Hence there is small abuse of notation when calling it a number of collisions.
%
Letting $\cP_M := \left\{ \boldsymbol{\mu} \in [0,1]^K : \mu_M^* > \mu_{M+1}^*\right\}$
be the set of bandit instances such that there is a strict gap between the $M$ best arms and the other arms, we now provide a regret decomposition for any $\boldsymbol{\mu} \in \cP_M$.

@@

Similarly to what we present above in Section~\ref{sec:2:lowerUpperBoundsRegret},
we use the Kullback-Leibler divergence to express our lower bound.
We remind that $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$ is the KL divergence between the Bernoulli distributions of means $x$ and $y$.
%
We first introduce the assumption under which we derive a regret lower bound, that generalizes the assumption of uniform efficiency given in Definition~\ref{def:2:uniformlyEfficientAlgorithm},
a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
% Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
% This class contains our proposals, \RandTopM{} and \MCTopM, but rules out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.

@@

Having a small regret on every problem instance, \ie, being uniformly efficient,
is a natural assumption,
that rules out algorithms tuned to perform well on specific instances only (\eg, fixed-armed algorithms).
%
From this assumption $\left(R_T(\boldsymbol{\mu},M,\cA)=o(T^\alpha)\right)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
${T}- \E_{\mu}[T_k(T)] {=} o(T^{\alpha})$, and so
\begin{equation}\label{eq:5:intermediate}
  % \vspace*{-5pt}  % XXX
  \sum_{j =1}^M\left(\frac{T}{M} - \E_{\mu}[T^j_k(T)]\right) = o(T^{\alpha}).
  % \vspace*{-5pt}  % XXX
\end{equation}

The additional assumption in \eqref{eq:5:SUE} further implies some notion of \emph{fairness}, as it suggests that each of the $M$ players spends on average the same amount of time on each of the $M$ best arms. Note that this assumption is satisfied by any strategy that is invariant under every permutation of the players, \ie, for which the distribution of the observations under $\cA_{\sigma} = (\cA_{\sigma(1)},\dots,\cA_{\sigma(M)})$ is independent from the choice of permutation $\sigma \in \Sigma_M$ of the set $\{1,\dots,M\}$.
In that case, it holds that  $\E_{\mu}[T_k^j(T)]=\E_{\mu}[T_k^{j'}(T)]$ for every arm $k$ and $(j,j') \in [M]$, hence \eqref{eq:5:SUE} and \eqref{eq:5:intermediate} are equivalent, and strong uniform efficiency is equivalent to standard uniform efficiency.
Note that the algorithms studied in Section~\ref{sec:5:algorithms} are permutation invariant, and \MCTopM-\klUCB{} is thus an example of strongly uniformly efficient algorithm, as we prove in Section~\ref{sec:5:upperbounds} that its regret is logarithmic on every instance $\mu \in \cP_M$.

% As we state it here, the constraint on fairness is new, and it was not studied as such in previous works.
In the proofs of the regret bounds for \rhoRand{} in \cite{Anandkumar11}, the authors briefly mention that their policy is invariant under permutation and that this yields a certain fairness guarantee, but without formalizing it more.
The notion of fairness defined above can be interpreted as a ``cooperative fairness'', opposed to the notion of ``arm fairness'' that has been studied in a few papers on single-player stochastic MAB.
For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of the total time steps, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K]$ is given and known to the algorithm.

@@

The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
Similarly to what we did in Section~\ref{sec:2:notations},
consider the observations $O_t^j$ that she gathered after $t$ rounds of its algorithm $\cA_j$ for a fixed player $j\in[M]$,
defined as $O_t^j := \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right)$,
where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).

Fix a problem $\bm{\mu}$,
% and as we do in Appendix~\ref{proof:5:BetterLowerBound},
then introduce an alternative model parametrized by $\bm{\lambda}$, with a small difference between $\bm{\mu}$.
Fix a sub-optimal arm $k$ and $\varepsilon>0$, and $\lambda_k = \mu^*_M + \varepsilon$ and $\lambda_{\ell} = \mu_{\ell}$ for any $\ell\neq k$.
We denote $\Pr_{\bm{\mu}}^{O_t^j}$ the distribution of the vector $O_t$ under the model $\bm{\mu}$ when using algorithm $\cA_j$.
Then we introduce the \emph{collision information term} as:
\begin{equation}\label{eq:5:collisionInformationTerm}
  \cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) := \sum_{t=1}^{T} \mathrm{KL}(\Pr_{\bm{\mu}}^{C_t | O_{t-1}}, \Pr_{\bm{\lambda}}^{C_t | O_{t-1}^j}).
\end{equation}

For a \emph{decentralized strategy} $\cA$ that has access to the sensing information (\ie, ruling out model \modeltrois), and satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) = \smallO{\log(T)}$,
we now state a problem-dependent asymptotic lower bound on the number of sub-optimal arms selections.
The additional hypothesis on $\cI_{\bm{\mu},\bm{\lambda}}(\cA,T)$ essentially says that ``the collisions do not bring too much information on the arm means'', as stated in \cite{KaufmannAbbas19}.
The theorem stated below is proven in the Appendix of \cite{Besson2018ALT},
% \ref{proof:5:BetterLowerBound},
and it also yields an asymptotic logarithmic lower bound on the regret.

\begin{theorem}\label{thm:5:BetterLowerBound}
\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
  Under observation models $\modelun$ and $\modeldeux$, consider a strongly uniformly efficient \emph{decentralized} policy $\cA=(\cA_1,\dots,\cA_M)$ and a problem $\boldsymbol{\mu}\in\cP_M$.
  Furthermore, if $\cA_j$ satisfies $\cI_{\bm{\mu},\bm{\lambda}}(\cA_j,T) = \smallO{\log(T)}$, then
  % \vspace*{-10pt}  % XXX
  \begin{equation}\label{eq:5:LBDraws}
    \forall j \in [M], \ \forall k \in \Mworst, \ \ \ \liminf_{T\to \infty} \frac{\E_{\mu}[T_k^j(T)]}{\log(T)} \geq \frac{1}{\kl(\mu_k, \mu_M^*)}.
  \end{equation}

  % \vspace{-0.3cm}  % XXX
  \noindent And so if $\cA_1,\dots,\cA_M$ all satisfy this hypothesis, from Lemma~\ref{lem:5:1stLowerBound}, it follows that
  % \vspace*{-5pt}  % XXX
  \begin{equation}\label{eq:5:ourLowerBound}
    \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log(T)}
    \geq M \times \left( \sum_{k \in \Mworst} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_M^*)} \right) .
  \end{equation}
\end{leftbar}  % XXX leftbar theorembar, comment if needed
\end{theorem}

Observe that the regret lower bound \eqref{eq:5:ourLowerBound} is tighter than the state-of-the-art lower bound in this setup
given by \cite{Zhao10}, that states that
\begin{equation}\label{eq:5:Zhao10LowerBound}
  \mathop{\lim\inf}\limits_{T \to +\infty} \frac{R_T(\boldsymbol{\mu}, M, \cA)}{\log(T)}
  \geq \sum_{k \in \Mworst} \left( \sum_{j=1}^{M} \frac{(\mu_M^* -  \mu_k)}{\kl(\mu_k, \mu_{j}^*)} \right),
\end{equation}
as for every $k \in \Mworst$ and $j \in [M]$, $\kl(\mu_k, \mu_j^*) \geq \kl(\mu_k, \mu_M^*)$
(see Figure~\ref{fig:5:CompLowerBounds} in Appendix~\ref{app:5:illustrationLowerBound}).
%
It is worth mentioning that \cite{Zhao10} considered the more general assumption for $\cA$ that there exists some numbers $(a_{k}^j)$ such that $a_{k}^j T - \E_{\mu}[T_k^j(T)] = o(T^\alpha)$ whereas in Definition~\ref{def:5:DecentralizedUniformEfficiency} we make the choice $a_{k}^j = 1/M$.
Our result could be extended to this case, but we chose to keep the notation simple and focus on \emph{fair allocation} of the optimal arms between players.

@@

Under observation model \modeltrois{} no sensing information is available and the previous algorithms cannot be used, as the sum of sensing information $S_k^j(t)$ and thus the empirical mean $\widehat{\mu}_k^j(t)$ cannot be computed, hence neither the indices $U_k^j(t)$. However, one can still define a notion of \emph{empirical reward} received from arm $k$ by player $j$, by introducing
%
% \vspace*{-5pt}
\begin{equation}
  \widetilde{S_k}^j(t) := \sum_{t=1}^T r^j(t) \indic(A^j(t) = k)
  \ \ \ \text{and letting} \ \ \ \widetilde{\mu_k}^j(t) := \widetilde{S_k}^j(t) \;/\; T_k^j(t).
\end{equation}

Note that $\widetilde{\mu_k}^j(t)$ is no longer meant to be an unbiased estimate of $\mu_k$ as it also takes into account the collision information, that is present in the reward. Based on this empirical reward, one can similarly defined modified indices as
%
\begin{equation}\label{eq:5:indexTilde}
  \widetilde{U_k}^j(t) := \begin{cases}
      \widetilde{\mu_k}^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widetilde{\mu_k}^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB.
  \end{cases}
\end{equation}

Given any of these two index policies (\UCB{} or \klUCB), the \Selfish{} algorithm is then playing like a single-player index policy (see Algorithm~\ref{algo:2:indexPolicy})
% \begin{equation}\label{algo:5:Selfish}
$A^j(t) \in \cU(\argmax_{ k \in [K]} \ \widetilde{U_k}^j(t-1))$.
% \end{equation}
The name ``selfish'' comes from the fact that each player is targeting, in a ``selfish'' way, the arm that has the highest index, instead of accepting to target only one of the $M$ best.
The reason that this may work precisely comes from the fact that $\widetilde{U_k}^j(t)$ is no longer an upper-confidence on $\mu_k$,
but some hybrid index that simultaneously increases when a transmission occurs and decreases when a collision occurs.
%
This behavior is easier to be understood for the case of \Selfish-\UCB{} in which, letting $N_k^{j,C}(t) = \sum_{s=1}^t \indic(C^j(t))$ be the number of collisions on arm $k$, one can show that the hybrid \Selfish{} index induces a penalty proportional to the fraction of collision on this arm and the quality of the arm itself:
\begin{equation}
  \widetilde{U_k}^j(t) = U_k^j(t) -
  \underbrace{\left(\frac{N_k^{j,C}(t)}{N_k^j(t)}\right)}_{\text{fraction of collisions}}
  \underbrace{\left(\frac{1}{N_k^{j,C}(t)}\sum_{t=1}^{T}Y_{A^j(t),t} \indic(C^j(t)) \indic(A^j(t) = k)\right)}_{\text{estimate of } \mu_k }.
\end{equation}

@@

We illustrate in Section~\ref{sub:5:SelfishFails} that \Selfish{} does have a (very) small probability to fail (badly), for some problem with small $K$,
which precludes the possibility of a logarithmic regret for any problem.
In most cases, it empirically performs similarly to all the algorithms described before,
and usually outperforms \rhoRand,
even if it neither exploits the sensing information, nor the knowledge of the number of players $M$.
%
Practitioners may still be interested by the algorithm, especially for Cognitive Radio applications in which sensing is hard or cannot be considered.
In fact, we used the \Selfish{} heuristic in Chapter~\ref{chapter:4}, in the different models, using \UCB{} or Thompson sampling as the underlying policy.
%
Thus we propose next our main contribution, the \MCTopM{} algorithm, proved to be asymptotically optimal for the identification of suboptimal arms (when using \klUCB), attaining order-optimal logarithmic regret, and outperforming all the other algorithms for the ``sensing case''.

@@

% In a single-player setting, \emph{index policies} are popular bandit algorithms: at each round one index is computed for each arm, that only depends on the history of plays of this arm and (possibly) some exogenous randomness. Then, the arm with highest index is selected. This class of algorithms includes the UCB family, in which the index of each arm is an Upper Confidence Bound for its mean, but also some Bayesian algorithms like Bayes-UCB \citep{Kaufmann12BUCB} or the randomized Thompson Sampling algorithm \citep{Thompson33,AgrawalGoyal11,Kaufmann12Thompson}.

% % concrete examples and how their are used within MPB

% \textbf{Index policies.}
% %
The approache we now describe for multi-player bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained. In particular, we focus on two types of indices:
\UCB{} \citep{Auer02}
and \klUCB{} \citep{KLUCBJournal}.
For player $j\in[M]$
%
denote $S_k^j(t) := \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained for arm $k$,
then $\widehat{\mu}_k^j(t) := S_k^j(t)/T_k^j(t)$ (if $T_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
\begin{equation}\label{eq:5:indexFor_UCB_klUCB}
  U_k^j(t) := \begin{cases}
      \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2T_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: T_k^j(t)\times \kl(\widehat{\mu}_k^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB,
  \end{cases}
\end{equation}
where $f(t)$ is some \emph{exploration function}, usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
A classical (single-player) UCB algorithm aims at the arm with largest index, as presented above in Section~\ref{sub:2:IndexPolicies}.
However, if each of the $M$ players selects the arm with largest UCB, all the players will end up colliding most of the time on the best arm.
To circumvent this problem, several coordination mechanisms have emerged, that rely on \emph{ordering} the indices and targeting \emph{one of} the $M$-best indices.

\paragraph{Two ideas for orthogonalization.}
\label{par:5:twoIdeasOrthogonalization}
%
On the one hand, the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices.
Even though some alternative without pre-agreement are proposed, they are quite complicated and we prefer to focus on other approaches.
%
On the other hand, the \rhoRand{} algorithm \citep{Anandkumar11} relies on randomly selected \emph{ranks}. %\footnote{\rhoRand{} only works if $M$ is known, and \cite{Anandkumar11} extended it to the case of an unknown $U$, with \rhoRandEst. Extending our algorithms for the case of unknown number of players $M$ is an interesting future work.}.
%
More formally, letting $\pi(k,\mathbf{g})$ be the index of the $k$-th largest entry in a vector $\mathbf{g}$,
% (for $k\in[K]$),
in \rhoRand{} each player maintains at time $t$ an internal rank $R^j(t)\in[M]$
and selects at time $t$,
% \begin{equation}
$A^j(t) := \pi\left(R^j(t), [U^j_\ell(t)]_{\ell=1,\dots,K}\right)$.
% \end{equation}
If a collision occurs, a new rank is drawn uniformly at random, $R^j(t+1) \sim \cU([M])$.
